---
title: "BLPestimatoR - Package for Demand Estimation"
author: "Daniel Brunner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BLPestimatoR - Package for Demand Estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
references:
- id: BLP1995
  title: Automobile Prices in Market Equilibrium
  author: 
  - family: Berry 
    given: Steven
  - family: Levinsohn 
    given: James
  - family: Pakes
    given: Ariel
  container-title: Econometrica 
  type: article-journal
  issued: 
    year: 1995
    
- id: Brunner2017
  title: Reliable Estimation of Random Coefficient Logit Demand Models 
  author: 
  - family: Brunner
    given: Daniel
  - family: Heiss
    given: Florian
  - family: Romahn
    given: Andre
  - family: Weiser
    given: Constantin
  container-title: DICE Discussion Paper No 267  
  type: article-journal
  issued: 
    year: 2017
    
- id: Nevo2001
  title: A Practitioner's Guide to Estimation of Random-Coefficients Logit Models of Demand
  author: 
  - family: Nevo
    given: Aviv
  container-title: Journal of Economics \& Management Strategy
  type: article-journal
  issued: 
    year: 2001
    
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Intro

`BLPestimatoR` provides an efficient estimation algorithm to perform the demand estimation described in -@BLP1995. The routine uses analytic gradients and offers a large number of optimization routines and implemented integration methods as discussed in -@Brunner2017.

This extended documentation demonstrates the steps of a typical demand estimation with the package:

- prepare the data with `BLP_data` (includes the specification of a model and providing integration draws for observed or unobserved heterogeneity)
- estimate the parameters with `estimate_BLP`
- showing the results with `summary`
- view the own- and crosspriceelasticities with `get_elasticities`

The use of the algorithm's building blocks is explained as well. The well-known training data for the cereal market from -@Nevo2001 is used for demonstration purposes.




# Data


## Model

Since version 0.1.6 the model is provided in R’s formula syntax and consists of five parts. The variable
to be explained is given by observed market shares. Explanatory variables are grouped into four (possibly
overlapping) categories separated by `|`:

 - linear variables
 - exogenous variables
 - random coefficients
 - instruments
 
Nevo's model can be translated into the following formula syntax:
  
```{r}
library(BLPestimatoR)

nevos_model <- as.formula("share ~  price + productdummy |
    0+ productdummy |
    price + sugar + mushy |
    0+ IV1 + IV2 + IV3 + IV4 + IV5 + IV6 + IV7 + IV8 + IV9 + IV10 + 
    IV11 + IV12 + IV13 + IV14 + IV15 + IV16 + IV17 + IV18 + IV19 + IV20")
```
 
The model is directly related to consumer $i$'s indirect utility from purchasing cereal $j$ in market $t$:

$$u_{ijt}=\sum_{m=1}^M x^{(m)}_{jt} \beta_{i,m}+\xi_{jt}+\epsilon_{ijt} \;\; \text{with}$$
$$\beta_{i,m}= \bar{\beta}_m  + \sum_{r=1}^R  \gamma_{m,r} d_{i,r}  +  \sigma_m \nu_{i,m}$$
and

- M = 4 random coefficients (`price`, `sugar`, `mushy` and an intercept)
- R = 4 demographics (`income`, `incomesq`, `age`, `child`)
- and the set of non-linear parameters to estimate:

$$\theta_2 = 
\begin{pmatrix}
\sigma_1 & \gamma_{1,1} & \cdots & \gamma_{1,R} \\
\sigma_2 & \gamma_{2,1} & \cdots & \gamma_{2,R} \\
\vdots & \vdots &  \ddots & \vdots \\
\sigma_M & \gamma_{M,1} & \cdots & \gamma_{M,R}
\end{pmatrix}$$



## Dataframe

Product related variables are collected in the dataframe `productData` with the following requirements:

 - missings are not allowed
 - character variables are automatically transformed to a set of dummy variables
 - a variable that describes market affiliation (`market_identifier`)
 
A variable that uniquely identifies a product in a market (`product_identifier`) is optional, but enhances clarity (interpreting elasticities, for example, is much easier). `market_identifier` and `product_identifier` together uniquely identify an observation, which is used by the function `update_BLP_data` to update any variable in the data (in this case `product_identifier` is mandatory).

In the cereal example, this gives the following dataframe:

```{r}
head(productData)
```


## Integration Draws
The arguments related to the numerical integration problem are of particular importance when providing
own integration draws and weights, which is most relevant for observed heterogeneity
(for unobserved heterogeneity, the straightforward approach is the use of automatic integration).

In the cereal data, both, observed and unobserved heterogeneity, is used for the random coefficients.
Starting with observed heterogeneity, user provided draws are collected in a list. Each list entry must be
named according to the name of a demographic. Each entry contains the following variables:

 - a variable `market_identifier` that matches each line to a market (same variable name as in `productData`)
 - integration draws for each market
 
In the cereal example, observed heterogeneity is provided as follows (list names correspond to the demographics):

```{r}
demographicData$income[1:4,1:5]

demographicData$incomesq[1:4,1:5]

demographicData$age[1:4,1:5]

demographicData$child[1:4,1:5]
```

If demographic input (`demographicData`) is missing, the estimation routine considers only coefficients for unobserved heterogeneity. This can be done by already implemented integration methods via `integration_method` as shown in the estimation section. In Nevo's cereal example however, a specific set of 20 draws is
given. For this situation, draws are also provided as a list (list names correspond to the formula's random coefficients and each list entry has a variable `market_identifier`):

```{r}
originalDraws$constant[1:4,1:5]

# renaming constants:
names(originalDraws)[1] <- "(Intercept)"

originalDraws$price[1:4,1:5]

originalDraws$sugar[1:4,1:5]

originalDraws$mushy[1:4,1:5]
```

As demonstrated above, list entries for draws of constants **must** be named `(Intercept)`. Other names of list entries must match the random coefficients specified in the formula.


## Calling BLP_data

Calling `BLP_data` structures and prepares the data for estimation and creates the data object:

```{r }
productData$startingGuessesDelta  <- c( log(w_guesses)) # include orig. draws in the product data

nevo_data <- BLP_data(model = nevos_model,
                      market_identifier="cdid",
                      par_delta = "startingGuessesDelta",
                      product_identifier = "product_id",
                      productData = productData,
                      demographic_draws = demographicData,
                      blp_inner_tol = 1e-6, blp_inner_maxit = 5000,
                      integration_draws = originalDraws, 
                      integration_weights= rep(1/20,20))
```


The arguments in greater detail:

- `model` provides the utility model as explained above

- `market_identifier` gives the name of the variable in `productData` that matches
each observation to a market

- `product_identifier` gives the name of the variable in `productData` that matches
each observation to a product (should be unique in a market)

- `productData` is given as a dataframe and `demographicData` as a list as described above

- `par_delta` gives the name of the variable in `productData` for mean utilities

- `blp_inner_tol` , `blp_inner_maxit`: arguments related to be BLP algorithm include the convergence threshold and the maximum number of iterations in the contraction mapping

- if integration draws are provided manually, `integration_draws` and
`integration_weights` need to be specified 

- for automatic integration the user specifies `integration_method`, for example
`integration_method= "MLHS"`, and the accuracy of the integration method by
`integration_accuracy` (for stochastic integration methods this equals the number
of draws)

If you decide to update your data later, you can use the function `update_BLP_data`. 



# Estimation 

## Starting guesses

The provided set of starting guesses `par_theta2` is matched with formula input and demographic data:

- rownames of `par_theta2` must match with the random coefficients specified in the formula (note: constants **must** be named `(Intercept)` ) 
- colnames of `par_theta2` must match with list entry names of `demographicData` and a column for unobserved heterogeneity (**must** be named `unobs_sd)
- `NA`s in `par_theta2` indicate the exclusion from estimation, i.e. the coefficient is assumed to be zero.
 
These requirements are demonstrated with Nevo's starting guesses:

```{r}
#before:
theta_guesses
theta_guesses[theta_guesses==0] <- NA 
colnames(theta_guesses) <- c("unobs_sd", "income", "incomesq", "age", "child" )
rownames(theta_guesses) <- c("(Intercept)", "price" , "sugar", "mushy")

# correctly named:
theta_guesses
```



## Calling estimateBLP

The following code performs the demand estimation:

```{r}

blp_est <- estimateBLP( blp_data=nevo_data,
                        par_theta2 = theta_guesses,
                        solver_method = "BFGS", solver_maxit = 1000, solver_reltol = 1e-6,
                        standardError = "heteroskedastic",
                        extremumCheck = FALSE ,
                        printLevel = 1 )

summary(blp_est)
```



The arguments in greater detail:

- `par_theta2` gives initial values for non-linear parameters to be optimized
over. Correct naming of columns and rows is important to allow correct matching. 

- `solver_method`, `solver_maxit` , `solver_reltol`: solver related arguments that specify the R internal optimization (`optim` function). Additional arguments can be passed to optim via `...`

- `standardError` can be specified as `homoskedastic`, `heteroskedastic` or
`cluster`. The latter requires the variable `group_structure` in `productData` giving the related
cluster.

- if `extremumCheck` is `TRUE`, numerical derivatives at the solver optimum are used
to check, if a local minimum was found

- `printLevel` controls for the amount of information that is provided during the
estimation


Many of these arguments have default values. In the following setting you see a
minimum of necessary arguments with an automatic generation of integration
draws and just unobserved heterogeneity. The summary output informs you about the most important default values.

```{r} 
nevo_data2 <- BLP_data(model = nevos_model,
                      market_identifier="cdid",
                     
                      product_identifier = "product_id",
                      productData = productData,
                      integration_method = "MLHS", 
                      integration_accuracy = 20, integration_seed = 213)

blp_est2 <- estimateBLP( blp_data=nevo_data2, printLevel = 1 )

summary(blp_est2) 
```

# Postestimation

## Standard Errors

Standard errors can be computed with three options that control for the unobserved characteristic $\xi$, which consists of $N$ elements. $\Omega$ denotes the variance covariance matrix of $\xi$. 

- option `homoskedastic` requires the standard deviation $\sigma_i$ for each $\xi_i \;\forall
i\in 1,\cdots,N$ to be identical: 
$$\Omega = \begin{pmatrix} \sigma & 0 &
\dots & 0\\ 0 & \sigma &  & 0\\ \vdots &  & \ddots & 0\\
0 & 0 & 0 & \sigma \\ \end{pmatrix}$$

- option `heteroskedastic` allows for individual standard deviations $\sigma_i$ for each
$\xi_i$ :
$$\Omega = \begin{pmatrix} \sigma_1 & 0 &
\dots & 0\\
0 & \sigma_2 &  & 0\\
\vdots &  & \ddots & 0\\
0 & 0 & 0 & \sigma_N \\ \end{pmatrix}$$


- option `cluster` allows for cluster individual variance covariance matrizes in each of $M$ cluster groups. For this option the argument `group_structure` needs to be specified in the function `BLP_data` to determine the cluster group. This gives the block-diagonal form with $\Sigma_m$ as the variance covariance matrix for all $\xi_i$ in cluster $m$:

$$\Omega = \begin{pmatrix} \Sigma_1 & 0 &
\dots & 0\\
0 & \Sigma_2 &  & 0\\
\vdots &  & \ddots & 0\\
0 & 0 & 0 & \Sigma_M \\ \end{pmatrix}$$




## Elasticities 

The following code demonstrates the calculation of elasticities. 

```{r}
get_elasticities(blp_data=nevo_data,
                 blp_estimation= blp_est,
                 variable = "price",
                 products = c("cereal_1","cereal_4"),
                 market = "market_2")
```

The value of the elasticity matrix in row $j$ and column $i$ for a variable $x$, gives the effect of a change in product $i$'s characteristic $x$ on the share of product $j$.

# Modular Examples

Further analysis like incorporating a supply side or performing a merger simulation often requires access to building blocks of the BLP algorithm. The following wrappers insure correct data inputs and access the internal functions of the algorithm.

In the following, you find an example of the contraction mapping and an evaluation of the GMM function at the starting guess:

```{r}

delta_eval <- getDelta_wrap(  blp_data=nevo_data,
                              par_theta2 = theta_guesses,
                              printLevel = 4 )


productData$startingGuessesDelta[1:6]
delta_eval$delta[1:6]
delta_eval$counter


gmm <- gmm_obj_wrap(  blp_data=nevo_data,
                      par_theta2 = theta_guesses,
                      printLevel = 2)

gmm$local_min
```
Printed distances in the contraction mapping are maximum absolute distance between the current vector of mean utilities and the previous one.


For any $\theta_2$, you can compute predicted shares: 
```{r}
shares <- getShares_wrap(  blp_data=nevo_data,
                           par_theta2 = theta_guesses, printLevel = 4)
shares[1:6]
```



The gradient contains two important building blocks as explained in the appendix of -@Nevo2001:

- $\frac{\partial s_{ijt}}{\partial \theta_2}$ , i.e. the derivative of individual $i$'s share of product $j$ in market $t$ with respect to non-linear parameters

- $\frac{\partial s_{ijt}}{\partial \delta}$ , i.e. the derivative of individual $i$'s share of product $j$ in market $t$ with respect to mean utilities

Both are used to compute the jacobian and are easy to obtain with the package as the following example demonstrates:

```{r}

derivatives1 <- dstdtheta_wrap(  blp_data=nevo_data,
                                 cm_input = delta_eval,
                                 market = "market_2")
derivatives2 <- dstddelta_wrap(  blp_data=nevo_data,
                                 cm_input = delta_eval,
                                 market = "market_2")

jacobian <- getJacobian_wrap(blp_data=nevo_data,
                             par_theta2 = theta_guesses,
                             printLevel = 2)

jacobian[1:5,1:4]

```

# References
